Biden Expected to Win 54% of Popular Vote in 2020 US Presidential Election

### 2020-11-02 outputs/paper/paper.Rmd
Able to knit.

```{r setup, include=FALSE}
library(haven)
library(tidyverse)
library(knitr)
library(palmerpenguins)
```
```{r}
#load required data and models

cleaned_data <- data.frame(read.csv(file = "../../inputs/cleaned_acs.csv"))

```
# Abstract
After a shocking upset in the 2016 US Presidential election, everyone has their eyes on the 2020 election to determine the leader of the free world for the next four years. In this Paper we trained a model using survey level survey results . Our model predicts that Joe Biden will win the 2020 Presidential election with a 53.5% of the popular vote and a confidence level of 95%. Our prediction as well as our breakdown of votes by demographic group could potentially provide the candidates of the election with information on how to target voters.

@@

When cleaning the data, we merged some of the factors of the variables to match the granularity of the data in the post-stratification dataset. This included splitting age responses into bins of size 10, reducing education to two bins ("High School or Less" and "Post Secondary or More"), and combining Asian Indian, Korean, Filipino, Vietnamese, and Pacific Islander ethnicities into "Asian (Other)" (Chinese and Japanese remained their own factors because this is the level of specificity available in the post-stratification dataset). Lastly, we took a subset of the dataset where respondents had decided to vote for either Trump or Biden in the 2020 US presidential election (for the purposes of being able to predict the popular vote using a binary model).
```{r, echo=FALSE}
reduced_data <- 
  raw_data %>% 
  select(vote_2020,
         age,
         gender,
         race_ethnicity,
         state,
         education)

@@

The distribution of each of our cleaned variables (with the exception of gender and state) are shown in the following two pages.

@@

```{r racedist, fig.height=4, fig.cap="Distribution of the race of respondents in percentages.", message=FALSE, echo=FALSE}
# summary table with counts for race
tibble_race <- data.frame(table(reduced_data$race_ethnicity)) %>%
  rename(
    race = Var1,
    count = Freq
  )
# distribution of sex as a bar graph in percentages
ggplot(tibble_race, aes(x = race, y = count/total_count)) + 
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0(round(count/total_count*100), "%")),
            color="white",
            position = position_stack(vjust = 0.5)) +
  xlab("Race") +
  ylab("Percentage of Respondents")
```

@@

Figure \@ref(fig:agedistribution) Out of the respondents, approximately 54% were female and 46% were male (Figure 2). Most of the respondents were in their 60's and 70's, while the next most common demographic were respondents in their 40's and 50's (Figure 1). This preliminary look at the dataset is fairly consistent with Canadian demographics according to the 2016 Census, with the female response being approximately 3% higher than expected and the average age being approximately 11 years older than expected (the average Canadian age is 41 while the average respondent age was 52). This older demographic makes sense as individuals less than 15 years of age were not eligible to respond to the survey and are therefore not represented here.

@@

```{r eval=FALSE}
# Plot Race
perc_race <- cleaned_data %>% count(race_ethnicity) %>% mutate(perc = n/nrow(cleaned_data))
race <- perc_race %>% ggplot(aes(x = race_ethnicity, y = perc)) + geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + labs(title = "Race/Ethnicity of Respondents in 2018 ACS data",
                                                                               x = "Race/Ethnicity", y = "Percentage", subtitle = "Figure X")
race

@@

# Plot state
perc_state <- cleaned_data %>% count(state) %>% mutate(perc = n/nrow(cleaned_data))
perc_state$state <- perc_state$state %>% factor(levels = sort(as.character.factor(perc_state$state))) 
state <- perc_state %>% ggplot(aes(x = state, y = perc)) + geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + labs(title = "State of Respondents in 2018 ACS data",
                                                                               x = "State", y = "Percentage", subtitle = "Figure X")
state

@@

@@

Our data is of penguins (Figure \@ref(fig:bills)).

```{r bills, fig.cap="Bills of penguins", echo = FALSE}
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)

```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=8, fig.height=4}

# This needs to be about the random data tha tI sasmpled?
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk way more about it. 
# Discussion

>>>>

```{r setup, include=FALSE}
library(haven)
library(tidyverse)
library(knitr)
library(palmerpenguins)
```
```{r loading}
# Load required data and models
cleaned_data <- data.frame(read.csv(file = "../../inputs/cleaned_acs.csv"))
```

# Abstract
After a shocking upset in the 2016 US Presidential election, everyone has their eyes on the 2020 election to determine the leader of the free world for the next four years. In this Paper we trained a model using survey level survey results . Our model predicts that Joe Biden will win the 2020 Presidential election with a 53.5% of the popular vote and a confidence level of 95%. Our prediction as well as our breakdown of votes by demographic group could potentially provide the candidates of the election with information on how to target voters.

@@

When cleaning the data, we merged some of the factors of the variables to match the granularity of the data in the post-stratification dataset. This included splitting age responses into bins of size 10, reducing education to two bins ("High School or Less" and "Post Secondary or More"), and combining Asian Indian, Korean, Filipino, Vietnamese, and Pacific Islander ethnicities into "Asian (Other)" (Chinese and Japanese remained their own factors because this is the level of specificity available in the post-stratification dataset). Lastly, we took a subset of the dataset where respondents had decided to vote for either Trump or Biden in the 2020 US presidential election (for the purposes of being able to predict the popular vote using a binary model).

```{r, echo=FALSE}
reduced_data <- 
  raw_data %>% 
  select(vote_2020,
         age,
         gender,
         race_ethnicity,
         state,
         education)

@@

The distributions of each of our cleaned variables (with the exception of gender and state) are shown in the following two pages.

@@

```{r racedist, fig.height=4, fig.cap="Distribution of the race of respondents in percentages.", message=FALSE, echo=FALSE}
# summary table with counts for race
tibble_race <- data.frame(table(reduced_data$race_ethnicity)) %>%
  rename(
    race = Var1,
    count = Freq
  )
tibble_race$race <- factor(tibble_race$race, 
                                      levels = c("White", "Black, or African American", 
                                                 "Asian (Chinese)", "Asian (Japanese)",
                                                 "Asian (Other)", "American Indian or Alaska Native", "Some other race"))
# distribution of sex as a bar graph in percentages
ggplot(tibble_race, aes(x = race, y = count/total_count)) + 
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0(round(count/total_count*100), "%")),
            color="black",
            position = position_stack(vjust = 1)) +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  xlab("Race") +
  ylab("Percentage of Respondents")
```

@@

Out of the respondents, approximately 50% were female and 50% were male. Most of the respondents were in their 60's and 70's, while the next most common demographic were respondents in their 40's and 50's (Figure 1). This preliminary look at the dataset is fairly consistent with Canadian demographics according to the 2016 Census, with the female response being approximately 3% higher than expected and the average age being approximately 11 years older than expected (the average Canadian age is 41 while the average respondent age was 52). This older demographic makes sense as individuals less than 15 years of age were not eligible to respond to the survey and are therefore not represented here.

@@

```{r eval=FALSE}
cleaned_data <- labelled::to_factor(cleaned_data)

# Plot Race
perc_race <- cleaned_data %>% count(race_ethnicity) %>% mutate(perc = n/nrow(cleaned_data))
race <- perc_race %>% ggplot(aes(x = race_ethnicity, y = perc)) + geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + labs(title = "Race/Ethnicity of Respondents in 2018 ACS data",
                                                                               x = "Race/Ethnicity", y = "Percentage", subtitle = "Figure X")
race

@@

# Plot state
perc_state <- cleaned_data %>% count(state) %>% mutate(perc = n/nrow(cleaned_data))
perc_state <- labelled::to_factor(perc_state)

perc_state$state <- perc_state$state %>% factor(levels = sort(as.character.factor(perc_state$state))) 
state <- perc_state %>% ggplot(aes(x = state, y = perc)) + geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + labs(title = "State of Respondents in 2018 ACS data", x = "State", y = "Percentage", subtitle = "Figure X")

state

@@

# Restore names for Biden and Trump votes
reduced_data$vote_2020 <- plyr::mapvalues(reduced_data$vote_biden,
                                           c(1, 0), 
                                           c("Joe Biden", "Donald Trump"))

@@

# Discussion
